Scenario 2 - MPX in the EU

### 2022-11-07 MS3.Rmd
Finalized table and data dictionary and uploading pdf

---
title: "Scenario 2 - MPX in the EU"
author: "Dawson Coblin and Jose Lazo"
date: '2022-11-07'
output: pdf_document
---

@@

This report aims to prepare data to analyze and assess the impact of MPV within  
each country of the EU, and understand how case rates may differ by region and 
other various demographic factors. These findings will aid the implementation
and appropriate response to the epidemic to local health departments.

@@

Milestone  | Description of Section
---------- | ----------------------
Milestone 1|Group Agreement
Milestone 2|Data-set Imported into R; Identify Key Data Elements and Types
Milestone 3|Clean Data-set Create Descriptive Statistics and Data Dictionary

Milestone 3 Sub-sections:  
1.Sub-setting Rows and Columns as Needed- Clean individual Data Frames  
2.Summarizing Data and Creating New Vectors, Organize Vectors for Use  
3.Clean Variables and Combined Data-sets  
4.Descriptive Stats of EU Data and Data Dictionary

@@

https://github.com/dcoblin/251.git

@@

a.	Team's preferred communication method

@@

\newpage
### 2) Milestone 2: Data-set Imported into R; Identify Key Data Elements and Types

```{r packages}
library(ggplot2)
library(tidyverse)
library(readr)
library(lubridate)
library(stringr)
library(readxl)
library(kableExtra)
```
```{r}
euro_mpx_cases <- read_csv("files/euro_mpx_cases.csv")
euro_census_stats <- read_csv("files/euro_census_stats.csv")
euro_pop_denominators <- read_csv("files/euro_pop_denominators.csv")
world_country_regions <- read_csv("files/world_country_regions.csv")
```
\newpage

@@

@@

euro_census_stats$`0-1000`[euro_census_stats$`0-1000` == 0] <- NA
euro_census_stats$`1000-9999`[euro_census_stats$`1000-9999` == 0] <- NA
euro_census_stats$`10000-99999`[euro_census_stats$`10000-99999` == 0] <- NA
euro_census_stats$`100000-199999`[euro_census_stats$`100000-199999` == 0] <- NA
euro_census_stats$`200000-499999`[euro_census_stats$`200000-499999` == 0] <- NA
euro_census_stats$`500000-999999`[euro_census_stats$`500000-999999` == 0] <- NA
euro_census_stats$GE1000000[euro_census_stats$GE1000000 == 0] <- NA
```

@@

@@

#Relocate the columns to a adjust fluidity of the data.
euro_mpx_cases<- relocate(euro_mpx_cases, floor_month, CountryCode, 
                          country_pop_2022, cases_grouped_monthly, 
                          .before = EU_region)
#Rename the sum of region cases that are confirmed.
euro_mpx_cases<- euro_mpx_cases %>%
   rename( sum_region_cases = `sum(cases_grouped_monthly)`)
```

@@

### Descriptive stats:
We find that the census information for 7 of the 29 countries in the EU are not
available. These "0" values were replaced with "NA" to prevent graphing issues.

### Region Data:
We see that the regions in EU include various numbers of countries, thus having 
unequal populations. A useful metric to create a graph will be the calculation 
of cases per 100,000. This will allow us to use 1 categorical variable (EU Region)
to a quantitative variable (number of confirmed cases)

@@

### The pertinent information per EU region is listed below:
```{r echo=FALSE}
Table_1 <- tibble(
  EU_Region = c("Eastern Europe", "Northern Europe", "Southern Europe",
               "Western Asia", "Western Europe" ),  
  Countries_per_Region = c(6, 9, 7, 1, 6),
  Total_Region_Pop = c(89171711, 37518701,133879004, 904705, 190251495),
  Total_cases_per_region = c( 217, 619, 7958, 4, 8214)
  )

Table_1
```
## Data Dictionary

```{r}
#Create the Data Dictionary

@@

kable(data_dic,
      col.names = c("Data Type", 
                    "Description"),
      caption = "Data Dictionary"
      ) 
```

>>>>

---
title: "Scenario 2 - MPX in the EU"
author: "Dawson Coblin and Jose Lazo"
date: '2022-11-07'
output: pdf_document
editor_options: 
  markdown: 
    wrap: 72
---

@@

This report aims to prepare data to analyze and assess the impact of MPV
within\
each country of the EU, and understand how case rates may differ by
region and other various demographic factors. These findings will aid
the implementation and appropriate response to the epidemic to local
health departments.

@@

| Milestone   | Description of Section                                           |
|-----------------------|-------------------------------------------------|
| Milestone 1 | Group Agreement                                                  |
| Milestone 2 | Data-set Imported into R; Identify Key Data Elements and Types   |
| Milestone 3 | Clean Data-set Create Descriptive Statistics and Data Dictionary |

Milestone 3 Sub-sections:\
1.Sub-setting Rows and Columns as Needed- Clean individual Data Frames\
2.Summarizing Data and Creating New Vectors, Organize Vectors for Use\
3.Clean Variables and Combined Data-sets\
4.Descriptive Stats of EU Data and Data Dictionary

@@

<https://github.com/dcoblin/251.git>

@@

```{=html}
<!-- -->
```
a.  Team's preferred communication method

@@

### 2) Milestone 2: Data-set Imported into R; Identify Key Data Elements and Types

```{r packages, include=FALSE}
library(ggplot2)
library(tidyverse)
library(readr)
library(lubridate)
library(stringr)
library(readxl)
library(kableExtra)
```

```{r}
euro_mpx_cases <- read_csv("files/euro_mpx_cases.csv")
euro_census_stats <- read_csv("files/euro_census_stats.csv")
euro_pop_denominators <- read_csv("files/euro_pop_denominators.csv")
world_country_regions <- read_csv("files/world_country_regions.csv")
```

@@

head(euro_pop_denominators,3)

@@

euro_census_stats$`0-1000`[euro_census_stats$`0-1000` == 0] <- NA
euro_census_stats$`1000-9999`[euro_census_stats$`1000-9999` == 0] <- NA
euro_census_stats$`10000-99999`[euro_census_stats$`10000-99999` == 0] <- NA
euro_census_stats$`100000-199999`[euro_census_stats$`100000-199999` == 0] <- NA
euro_census_stats$`200000-499999`[euro_census_stats$`200000-499999` == 0] <- NA
euro_census_stats$`500000-999999`[euro_census_stats$`500000-999999` == 0] <- NA
euro_census_stats$GE1000000[euro_census_stats$GE1000000 == 0] <- NA

head(euro_census_stats, 3)
```

@@

head(euro_mpx_cases,3)
```

@@

#Relocate the columns to a adjust fluidity of the data.
euro_mpx_cases<- relocate(euro_mpx_cases, floor_month, CountryCode, 
                          country_pop_2022, cases_grouped_monthly, 
                          .before = EU_region)
#Rename the sum of region cases that are confirmed.
euro_mpx_cases<- euro_mpx_cases %>%
   rename( sum_region_cases = `sum(cases_grouped_monthly)`)

head(euro_mpx_cases,3)
```

@@

### Descriptive stats:

We find that the census information for 7 of the 29 countries in the EU
are not available. These "0" values were replaced with "NA" to prevent
graphing issues.

### Region Data:

We see that the regions in EU include various numbers of countries, thus
having unequal populations. A useful metric to create a graph will be
the calculation of cases per 100,000. This will allow us to use 1
categorical variable (EU Region) to a quantitative variable (number of
confirmed cases)

@@

### The pertinent information per EU region is listed below:

```{r, include=FALSE}

Table_1.1 <- euro_mpx_cases %>% group_by(EU_region, country_pop_2022, sum_region_cases) %>%
  count(CountryCode) %>%
  select(CountryCode,EU_region,country_pop_2022,sum_region_cases) %>%
  ungroup() %>%
  group_by(EU_region, sum_region_cases) %>%
  summarize(Total_region_pop = sum(country_pop_2022))
Countries_per_Region = c(6, 9, 7, 1, 6)
Table_1.1 <- as.data.frame(cbind(Table_1.1, Countries_per_Region))
```

```{r, echo=FALSE}
kable(Table_1.1,
      col.names = c("EU Region", 
                    "Total Cases Per Region",
                    "Total Region Population",
                    "Countries Per Region"),
      caption = "Monkeypox Cases Per Region"
      ) %>%
  kable_styling(latex_options = "HOLD_position")

## Data Dictionary

```{r, echo=FALSE}
#Create the Data Dictionary

@@

kable(data_dic,
      col.names = c("Data Type", 
                    "Description"),
      caption = "Data Dictionary"
      ) %>%
  kable_styling(latex_options = "HOLD_position")
```

```{r, include=FALSE}
#euro_mpx_cases %>% 
#  ggplot(., aes(x = EU_region, y = country_pop_2022, fill = EU_region)) +
#  geom_bar(stat = "identity", position = position_dodge()) +
#  theme_minimal() +
#  labs(title = "Total Population 2022 by Region", x = "", y = "Pop") +
#  theme(axis.title = element_text(size = 14),
#        axis.text = element_text(size = 8),
#        axis.ticks.x = element_blank(),
#        legend.title = element_blank(),
#        legend.position = "none")
```
